---
title: "Tokenisation and stopwords"
teaching: 0
exercises: 0

---

:::::::::::::::::::::::::::::::::::::: questions 

- "How do we prepare text for analysis and measure the sentiment of the text?"

::::::::::::::::::::::::::::::::::::::::::::::::

::::::::::::::::::::::::::::::::::::: objectives

- "Using specific packages to perform text preparation and sentiment analysis"
::::::::::::::::::::::::::::::::::::::::::::::::

```{r, include = FALSE}
library(tidyverse)
library(tidytext)
library(tm)
articles <- read_csv("data/articles.csv", na = c("NA", "NULL", ""))
```

## Tokenisation
Since we are working with text mining we focus on the `text` coloumn. We do this because the coloumn contains the text from articles.

To tokenise a coloumn, we use the functions `unnest_tokens()` from the `tidytext`-package. The function gets two arguments. The first one is `word`. This defines that the text should be split up by words. The second argument, `text`, defines the column that we want to tokenise.

```{r}
articles_tidy <- articles %>% 
  unnest_tokens(word, text)
```

:::: callout

### Tokenisation

The result is 118,269 rows. The reason is that the `text`-column is replaced by a new column named `word`. This columns contains all words found in all of the articles. The information from the remaining columns are kept. This makes is possible to dermine which article each word belongs to.

::::::

## Stopwords
The next step is to remove stopwords. We have chosen to use the stopword list from the package `tidytext`. The list contains 1,149 words that are considered stopwords. Other lists are available, and they differ in terms of how many words they contain.

```{r}
data(stop_words)
stop_words
```

:::: callout

### Adding and removing stopwords

You may find yourself in need of either adding or removing words from the stopwords list.

Here is how you add and remove stopwords to a predefined list.

**Add stopwords**
First, create a tibble with the word you wish to add to the stop words list

```{r}
new_stop_words <- tibble(
  word = c("cat", "dog"),
  lexicon = "my_stopwords"
)
```

Then make a new stopwords tibble based on the original on, but with the new words added.

```{r}
updated_stop_words <- stop_words %>%
  bind_rows(new_stop_words)
```

Run the following code to see that the added lexicon `my_stopwords` contains two words.
```{r}
updated_stop_words %>% 
  count(lexicon)
```



**remove stopword**
First,  create a vector with the word(s) you wish to remove from the stopwords list
 
```{r}
words_to_remove <- c("cat", "dog")
```

Then remove the rows containing the unwanted words.
```{r}
updated_stop_words <- stop_words %>%
  filter(!word %in% words_to_remove)
```

Run the following code to see that the added lexicon `my_stopwords` nolonger exists.
```{r}
updated_stop_words %>% 
  count(lexicon)
```



::::::

In order to remove the stopwords from `articles_tidy`, we have to use the `anti_join`-function. 

```{r}
articles_anti_join <- articles_tidy %>% 
  anti_join(stop_words, by = "word")
```

:::: callout

### `Join` and `anti_join`




::::::

## Frequency analysis

A word frequency is a relatively simple analysis. It meassures how often words occur in a text. 


```{r}
articles_anti_join %>% 
  count(word, sort = TRUE)
```

The previous code chunk resulted in a list containing the most frequent words. The words are from articles about both presidents, and they are sorted based on frequency with the highest number on top.

A closer look at the list may reveal that some words are irrelevant. Given that the articles in the dataset are about the two presidents' respective inaugurations, we consider the words below irrelevant for our analysis. Therefore, we make a new dataset without those words.

```{r}
articles_filtered <- articles_anti_join %>%
  filter(!word %in% c("trump", "trump’s", "obama", "obama's", "inauguration", "president"))

articles_filtered %>% 
  count(word, sort = TRUE)
```
The words deemed irrelevant are no longer on the list above.

Instead of a general list it may be more interesting to focus on the most frequent words belonging to articles about the two presidents respectively.

```{r}
articles_filtered %>%
  count(president, word, sort = TRUE)
```
It can be a bit tricky to keep an overview of the words associated with each president. For instance, the people is associated with both presidents. This is easy to see, as the two words are right next to each other. America, however, are further apart, although this word i also associated with the presidents. To that end, a visualisation may help.


```{r}
articles_filtered %>%
  count(president, word, sort = TRUE) %>% 
  group_by(president) %>%
  slice(1:10) %>% 
  ggplot(mapping = aes(x = n, y = word, colour = president, shape = president)) +
  geom_point() 
```
The plot above shows the top-ten words associated Obamma and Trump respectively. If a word features on both presidents' top-ten list, it only occures once in the plot. This is why the plot doesn't contain 20 words in total.

Another interesting aspect to look at would be the most frequent words used in relation to each president. In this analysis the president is the guiding principle.

```{r}
articles_filtered %>%
  count(president, word, sort = TRUE) %>% 
  pivot_wider(
    names_from = president,
    values_from = n
  )
```

```{r}
articles_filtered %>%
  group_by(president) %>% 
  count(word, sort = TRUE) %>% 
  top_n(10) %>% 
  ungroup() %>% 
  mutate(word = reorder_within(word, n, president)) %>% 
  ggplot(aes(n, word, fill = president)) +
  geom_col() +
  facet_wrap(~president, scales = "free") +
  scale_y_reordered() + 
  labs(x = "word occurrences")

```
The analyses just made can easily be ajusted. For instance, if we want look at the words by `pillar_name` instead of by `president`, we just need to replace `president` with `pillar_name` in the code.

```{r}
articles_filtered %>%
  count(pillar_name, word, sort = TRUE) %>% 
  group_by(pillar_name) %>%
  slice(1:10) %>% 
  ggplot(mapping = aes(x = n, y = word, colour = pillar_name, shape = pillar_name)) +
  geom_point() 
```










::::::::::::::::::::::::::::::::::::: keypoints 

- "All natural language texts must be prepared for analysis"

::::::::::::::::::::::::::::::::::::::::::::::::
